---
title: "Learning Base R"
author: "Michael Kalata"
date: "November 22, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


```


# Preface

R is a free software package that includes a rich set of data structures, computational functions, graphical functions, statistical procedures, programming capability, and much more. 


```{r}
# Different histograms based on number of random variates
hist(rnorm(134))
hist(rnorm(10))
hist(rnorm(100))
hist(rnorm(1000))
hist(rnorm(10000))

```

The greater-than symbol (>) at the beginning of the line is a prompt that awaits an R command from the user. Code that you want to run but not save can be ran in the console tab below.

The function name "hist" is short for "histogram." 

```{r}
barplot(VADeaths, beside = TRUE, legend = TRUE)
VADeaths
help(VADeaths)
```


# Chapter 2: R as Calculator

# 2.1 Order of operations
```{r}

2 + 9 * 4            # PEMDAS; place spaces around operators
# [1] that appears before 38 indicates the result is a vector of length 1

2+9*4               # no spaces around operators; it still works

4 + 3 / 10 ^2       # exponential first, then division, then addition

# Parantheses used to alter order of operations
4 + (3 / 10) ^ 2    # division first, then exponentiation, then addition

(4 + 3) / 10 ^ 2    # addition first, then exponentiation, then diviion

(4 + 4 / 10) ^ 2    # division first, then addition, then exponentiation

((4 + 3) / 10) ^ 2  # addition first, then division, then exponentiation

2 * 3 -             # + prompt for more input (when run in console)
+ 7                 # completed command 2 * 3 - 7

1 / 4               # the display suppresses trailing zeros

1 / 3               # default seven dispaly digits

8 * 10 ^ 3          # exponentiation first

1111111 * 1111111   # no commas on constants; result in scientific notation


# 2.2 Number of digits displayed
pi                  # built in constant,

options(digits = 10)# display 10 digits
pi

options(digits = 5)# display 5 digits
pi

options(digits = 22)# display 22 digits
pi

1111111 * 1111111   # now exact

789 ^ 10            # expressed in scientific notation

17.046875           # stored exactly

6.02214e23          # stored dependent on machine; book shows full number

1 / 17 ^ 7          # a small number


# 2.3 Undefined calculations

1/0

Inf - Inf

-1 / 0

0 / 0

(-9) ^ (1 / 2)
```

# Chapter 3: Simple Object

We now define an object in R. An object is typically called a variable in most computer languages, but the designer of R chose the term object in order to highlight their generality as a data structure. 

The = and <- assignment operators assign a numeric value to an object, The two assignment operators are equivalent. The latter has been defined to look a bit like an arrow that points to the left. (No space is allowed between the < and the - when using the arrow-like version of the assignment operator.)
```{r}
# 3.1 Assigning numeric values to objects 

x = 2


```
